---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
download.file("https://raw.githubusercontent.com/vfoss/RepData_PeerAssessment1/master/activity.csv", 
              "activity.csv", method="curl")

activity <- read.csv("activity.csv", colClasses=c("numeric","Date","numeric"),
                     row.names = NULL, stringsAsFactors = FALSE, header=TRUE)
```

## What is the total number of steps taken per day?

1. Calculate the total number of steps taken per day
```{r}
allsteps <- tapply(activity$steps, activity$date, sum, na.rm=TRUE)
```
  
2. Make a histogram of the total number of steps taken each day
```{r}
hist(allsteps, xlab="Total Number of Steps taken per Day")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(allsteps)
```

```{r}
median(allsteps)
```

## What is the average daily activity pattern?
1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepintervals <- tapply(activity$steps,activity$interval,mean,na.rm=TRUE)

plot(names(stepintervals), stepintervals, type="l", 
           xlab="5-minute Interval", ylab="Average Number of Steps Taken")
```
  
2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
as.numeric(names(stepintervals)[which.max(stepintervals)])
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
